---
title: "5320HW1"
author: "Daniel Silva"
date: "2/17/22"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1 

```{r}
set.seed(5320)
n <- 10
n2 <- 100000
x1 <- rnorm(n,8,4)
x2 <- runif(n,0,12)
X <- rbind(x1,x2)
ex1 <- rnorm(n2,8,4)
ex2 <- runif(n2,0,12)
eX <- rbind(x1,x2)
xtest <- cbind(x1,x2)
cov(xtest)
```


# Part a

Find the expectation E[X] and the variance matrix Var[X] of the random vector X

$E[X] = [8,6]'$

$Var[X] = [16,0;0,12]$

```{r}
mu_x = matrix(c(8,6),2,1,byrow=TRUE)
var_x = matrix(c(16,0,0,12),2,2,byrow=TRUE)
```

# Part b 

Find a matrix A and a column vector b such that Y = [Y1,Y2,Y3]' = AX+b

$[1,1,0;1,-1,0;5,0,0]*[x1,x2,x3]' = A*[x1,x2,x3]' + b$

where $A = [1,1,0;1,-1,0;5,0,0]$ and $b = [0,0,10]'$

```{r}
y1 = x1+x2
y2 = x1-x2
y3 = 5*x1+10
Y = rbind(y1,y2,y3)

ey1 = ex1+ex2
ey2 = ex1-ex2
ey3 = 5*ex1+10
eY = rbind(ey1,ey2,ey3)


A = matrix(c(1,1,0,1,-1,0,5,0,0),3,3,byrow=TRUE)
b = matrix(c(0,0,10),3,1,byrow=TRUE)
 
```

# Part c 

Find the expectation E[Y] and Var[Y] of the random variable Y. 

$E[Y] = E[[Y1,Y2,Y3]'] = [E[Y1],E[Y2],E[Y3]]'$ 
$= [ E[X1]+E[X2],E[X1]-E[X2],5*E[X1]+10]'$
$= [ 8+6, 8-6, 5*8 +10 ]'$
$= [14, 2, 50]'$

$Var(Y) = Var([Y1,Y2,Y3]')$

$$Var(Y) = \begin{bmatrix} Var(Y1) & Cov[Y1,Y2] & Cov[Y1,Y3] \\
 Cov[Y2,Y1] & Var(Y2) & Cov[Y2,Y3] \\
Cov[Y3,Y1] & Cov[Y3,Y2] & Var(Y3) \\
\end{bmatrix}$$


$$Var(Y) = \begin{bmatrix} Var(X1+X2) & Cov[X1+X2,X1-X2] & Cov[X1+X2,5*X1+10] \\
Cov[X1-X2,X1+X2] & Var(X1-X2) & Cov[X1-X2,5*X1+10] \\
Cov[5*X1+10,X1+X2] & Cov[5*X1+10,X1-X2] & Var(5*X1+10) \\
\end{bmatrix}$$

$$Var(Y) = \begin{bmatrix} var(x1)+var(x2) & var(x1)-var(x2) & 5*var(x1) \\
var(x1)-var(x2) & Var(x1)-var(x2) & 5*var(x1) \\
5*var(x1) & 5*var(x1) & 25*var(x1)+10 \\
\end{bmatrix}$$

$$Var(Y) = \begin{bmatrix} 28 & 4 & 80 \\
4 & 4 & 80 \\
80 & 80 & 310 \\
\end{bmatrix}$$
# Part d

Simulations n = 10

```{r}
m = cbind(y1,y2,y3)
colMeans(m)
cov(m)
```

# Part e 

Simulations n = 100,000

```{r}
em = cbind(ey1,ey2,ey3)
colMeans(em)
cov(em)
```

## Problem 2

Seber Lee 1a.1

Prove that if a is a vector of constants with the same dimension as the
random vector X ,., then

$E[(X-a)(X-a)']=Var[X]+(E[X]-a)(E[X]-a)'$

Proof:

Let 

$E[X] = \mu$


$E[(X-a)(X-a)'] = E[(X-\mu+\mu -a)(X-\mu+\mu -a)'] \\$
$= E[(X-\mu)(X=`-\mu)' + (X-\mu)(\mu - a)' + (\mu -a)(X- \mu)' + (\mu - a)(\mu - a)'] \\$
$= E[(X-\mu)(X=`-\mu)']+E[(X-\mu)(\mu - a)']+E[(\mu -a)(X- \mu)']+E[(\mu - a)(\mu - a)'] \\$
$= Var(X) + (E[X] - a)(E[X]-a)'$



## Problem 3

```{r}
sigma_X = matrix(c(1,1,3/4,1,2,3,3/4,3,3),3,3,byrow=TRUE)
B = matrix(c(2,-1,1),1,3,byrow=TRUE)
B%*%sigma_X%*%t(B)

B%*%matrix(c(1,1,-1),3,1,byrow=TRUE)

B%*%sigma_X%*%t(B)
```
































